---
title: "Disrete Bernoulli"
output: html_notebook
---


# Simulations for the discrete Bernoulli experiment


```{r setup}
library(SBC)
library(tidyverse)
library(patchwork)

library(future)
plan(multisession)
options(SBC.min_chunk_size = 200)
```


Generate datasets + setup a set of different generated quantities.

I create a new simple class of SBC backends that just use one function to generate samples when $y = 0$ and another when $y = 1$.

```{r}
set.seed(5465822)

N_sims_disc_gq <- 1000
N_sims_disc_gq_large <- 50000

theta_options <- c(1/3, 2/3)

variables_disc_gq <- sample(theta_options, size = N_sims_disc_gq_large, replace = TRUE)
generated_disc_gq <- 
  purrr::map(variables_disc_gq, ~ list(y = rbinom(1, size = 1, .x)))

datasets_disc_gq_large <- SBC_datasets(variables = posterior::draws_matrix(theta = variables_disc_gq),
                               generated = generated_disc_gq)

datasets_disc_gq <- datasets_disc_gq_large[1 : N_sims_disc_gq]


disc_backend_func <- function(prob0, prob1, theta_opts = theta_options, N_draws = 100) {
  if(length(prob0) != length(theta_options) - 1 ||
     length(prob1) != length(theta_options) - 1) {
    stop("Incorrect number of probabilities")
  }
  structure(list(prob0 = prob0, prob1 = prob1, 
                 theta_options = theta_opts, N_draws = N_draws), 
            class = "disc_backend_func")
}

SBC_fit.disc_backend_func <- function(backend, generated, cores) {
  if(generated$y == 0) {
    prob_all <- c(backend$prob0, 1 - sum(backend$prob0))
  } else if (generated$y == 1) {
    prob_all <- c(backend$prob1, 1 - sum(backend$prob1))
  } else {
    stop("Invalid")
  }
  posterior::draws_matrix(theta =  sample(backend$theta_options, size = backend$N_draws, prob = prob_all, replace = TRUE))
}

SBC_backend_iid_draws.disc_backend_func <- function(backend) {
  TRUE
}

my_globals <- c("SBC_fit.disc_backend_func", "SBC_backend_iid_draws.disc_backend_func")

gq_disc_gq <- generated_quantities(
   log_lik = dbinom(y, size = 1, prob = theta, log = TRUE),
   sq = (theta - 0.5) ^ 2,
   sin3_2 = sin(3/2 * pi * theta),
   abs = abs(theta - 1/2),
   saw = ifelse(theta < 1/2, theta, -1/2 + theta),
   swap = ifelse(theta < 1/2, theta, theta - 1),
   saw_quad = ifelse(theta < 1/2, theta^2, -1/2 + theta^3))

```

## Correct posterior

Setup a backend using the correct analytic posterior - it passes SBC including all GQs

```{r}
backend_disc_gq_ok <- disc_backend_func(
  prob0 = (1 - theta_options[1])/sum(1- theta_options),
  prob1 = theta_options[1] / sum(theta_options))

res_disc_gq_ok <- compute_SBC(datasets_disc_gq_large, backend_disc_gq_ok, keep_fits = FALSE,
                        gen_quants = gq_disc_gq, globals = my_globals)

plot_rank_hist(res_disc_gq_ok)
plot_ecdf_diff(res_disc_gq_ok)
```


## Understandng problems

```{r}
backend_disc_gq_test <- disc_backend_func(
  # prob0 = 0.1,
  # prob1 = 0.1)
  prob0 = 0.5 * (1/2 + 2/3),
  prob1 = 0.5 * (1/2 + 1/3))

res_disc_gq_test <- compute_SBC(datasets_disc_gq_large, backend_disc_gq_test, keep_fits = FALSE,
                        gen_quants = gq_disc_gq, globals = my_globals)

plot_rank_hist(res_disc_gq_test)
plot_ecdf_diff(res_disc_gq_test)
```
