---
title: "R Notebook"
output: html_notebook
---

Note on notation: we will denote the integral of $f(x)$ w.r.t $x$ over domain $X$ as $\int_X \mathrm{d}x \: f(x)$.

$$
\pi_{\text{joint}}(y, \theta) = \pi_{\text{obs}}(y | \theta) \pi_\text{prior}(\theta)\\
\pi_\text{marg}\left(y \right) = \int_\Theta \mathrm{d} \theta \: \pi_{\text{obs}}(y | \theta) \pi_\text{prior}(\theta)\\
\pi_{\text{post}}(\theta | y) = \frac{\pi_{\text{obs}}(y | \theta) \pi_\text{prior}(\theta)}{\pi_\text{marg}\left(y \right)}
$$

Alternate $q$ def:

Given a model, fitting algorithm and a function $f$, we define fitted pushforward CDF

$$
C_{\phi,f}(s | y) = \int_{\Theta}\mathrm{d}\theta\mathbb{\,I}\left[f\left(\theta, y \right) \leq s\right]\phi\left(\theta | y\right)
$$ 

and true pushforward CDF 

$$
C_{f}\left( s | y\right)=\int_{\Theta}\mathrm{d}\theta\mathbb{\,I}\left[f\left(\theta, y \right) \leq s\right]\pi_{\text{post}}(\theta | y)  \\


$$

Where $\mathbb{I}$ is the indicator function

$$
D_{\phi,f}(s | y) = \int \mathrm{d}\theta \: \phi(\theta | y) \mathbb{I}\left[ f(\theta, y) = s \right]
$$


For fixed $\tilde\theta$ and $y$ we define the random variable 
$\rho_{\phi,f}(\tilde\theta | y)$ (continuous rank) as

$$
\theta \sim \phi_y \\
e \sim \text{Uniform}_c(0, 1)\\
\rho_{\phi,f}(\tilde\theta | y) = \mathbb{P} \left(f(\theta, y) < f\left(\tilde\theta,  y \right) \right) + e \mathbb{P} \left(f(\theta, y) = f \left(\tilde\theta, y \right) \right) =\\
C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) - e D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)
$$


$$
\mathbb{P}\left(\rho_{\phi,f}\left(\tilde\theta | y \right) \leq x | \tilde\theta \right) = 
\begin{cases}
  0 & \text{if } x < C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) - D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) \\
  1 - \frac{C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) - x}{D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)} & 
    \text{if } D_{...} > 0 \text{ and } C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) - D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) \leq x < C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) \\
  1 & \text{if } x \geq C_{\phi,f} \left(f \left(\tilde\theta, y \right) \right) 
\end{cases} =
\\
= \begin{cases}
  0 & \text{if } f(\tilde\theta, y) > \inf \left\{s : x \leq C_{\phi,f}(s | y) -  D_{\phi,f} \left(s | y \right) \right\} \\
  1 - \frac{C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) - x}{D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)} & 
     \text{if } D_{...} > 0 \text{ and } f(\tilde\theta, y) = \inf\{s : x \leq C_{\phi,f}(s | y) \} \\
  1 & \text{if } f(\tilde\theta, y) \leq \inf\{s : x \leq C_{\phi,f}(s | y)\}
\end{cases}
$$

TODO: figure out exact ineqaulities above

```{r}
x <- 0.2
f <- 1

c_phi <- function(x) { pbinom(x, size = 2, prob = 0.5) }

inv_c_phi <- function(x) { qbinom(x, size = 2, prob = 0.5) }

inv_c_m_d_phi <- function(x) { }


```


```{r}
c_phi <- function(x) { 
  (floor(x)  + 1) / 3
#   dplyr::case_when(
#   x < 0 ~ 0,
#   x < 1 ~ 1/3,
#   x < 2 ~ 2/3,
#   TRUE ~ 1
# ) 
}

inv_c_phi <- function(x) {
  #inf s : x <= c_phi(s)
  dplyr::case_when(
    #x == 0 ~ 1,
    x < 1/3 ~ 0,
  x < 2 ~ 1/3,
  x < 3 ~ 2/3,
  TRUE ~ 1
) 
}
```


x = 0.5
f(tilde t) = 1 
phi === bernoulli(0.5)
P(rho <= x) =  0 ; x <= 0.5
           =  (x - 0.5) * 2 ; 1 > x > 0.5
           =  1 ; x >= 1

ALT:
P(rho <= x | tilde t) = 0 ; f(tilde t)


$$
x \leq C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) \\
f(\tilde\theta,y) \geq \{s: C_{\phi,f} \left(s | y \right) \geq x  \}
\\
a
\\
C^{-1}_{\phi,f}(x | y) \leq C^{-1}_{\phi,f} \left ( C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) |y \right) \\
$$

$$
q_{\phi,f}(x|y) = 
\mathbb{P}(\rho_{\phi,f}(\tilde\theta | y) \leq x) = 
\int \mathrm{d} \tilde\theta \: \pi_\text{post}(\tilde\theta | y) \mathbb{P}(\rho_{\phi,f}(\tilde\theta | y) \leq x) = \\

$$

With no ties:

$$
q_{\phi,f}(x|y) = 
\int \mathrm{d} \tilde\theta \: \pi_\text{post}(\tilde\theta | y) \mathbb{I}\left[C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) \leq x  \right] = \\
\int \mathrm{d} \tilde\theta \: \pi_\text{post}(\tilde\theta | y) \mathbb{I}\left[f \left(\tilde\theta, y \right) \leq C^{-1}_{\phi,f}(x | y)\right] =  C_f \left(C^{-1}_{\phi,f}(x | y) | y \right)
$$

Trying to prove that continuous SBC implies sample SBC with ties

$$
\text{rank_lower}(\tilde\theta |y) = \mathbb{P}(f(\theta, y) < f(\tilde\theta, y)) + \frac{1}{2} \mathbb{P}(f(\theta, y) = f(\tilde\theta, y)) = \\
\int_\Theta \mathrm{d}\theta \: \left(\mathbb{P}(f(\theta, y) < f(\tilde\theta, y) \right)
$$


$$
P(N_\text{less} = n, N_\text{equals} = k | \tilde\theta, y) = \frac{M!}{n!k!(M - k - n)!} \left(\mathbb{P}(f(\theta, y) < f(\tilde\theta, y))\right)^n \left(\mathbb{P}(f(\theta, y) = f(\tilde\theta, y))\right)^k\left(\mathbb{P}(f(\theta, y) > f(\tilde\theta, y))\right)^{M-k-n} \\

P(N_\text{less} + K = r | \tilde\theta, y) = \sum_{n = 0}^r\sum_{k = r - n}^M \frac{1}{k + 1}P(N_\text{less} = n, N_\text{equals} = k | \tilde\theta) \\
$$

$$
\mathbb{P} \left( f(y, \theta_1) < f(y, \tilde \theta) \land \ldots \land f(y, \theta_n) < f(y, \tilde \theta) \land \right.\\
  \land f(y, \theta_{n + 1}) = f(y, \tilde \theta) \land \ldots \land f(y, \theta_{n + n_\text{eq}}) = f(y, \tilde \theta)  \\
  \left. \land f(y, \theta_{n + n_\text{eq} + 1}) > f(y, \tilde \theta) \land \ldots \land f(y, \theta_{M}) > f(y, \tilde \theta) \land K = k | y \right) = \\
  \frac{1}{n_\text{eq}}
$$



$$
P(N_\text{less} + K = r | y) = \int_0^1 \mathrm{d}x \: {M \choose r}\left(q^{-1}(x|y)\right)^r \left(1 - q^{-1}(x|y)\right)^{(M - r)} ?? \\
P\left(\rho_{\phi,f}(\tilde\theta|y) \leq \frac{r + 1}{M + 1}\right) ?? \\

\int_\Theta \mathrm{d}\tilde\theta \int_Y \mathrm{d}y \: P\left(\frac{r}{M + 1} < \rho_{\phi,f}(\tilde\theta|y) \leq \frac{r + 1}{M + 1}\right) = 
\int_\Theta \mathrm{d}\tilde\theta \int_Y \mathrm{d}y \left(q\left(\frac{r + 1}{M + 1} |y \right) - q\left(\frac{r}{M + 1} |y \right) \right) = \\
\int_\Theta \mathrm{d}\tilde\theta \int_Y \mathrm{d}y \: q\left(\frac{r + 1}{M + 1} |y \right) - \int_\Theta \mathrm{d}\tilde\theta \int_Y \mathrm{d}y \: q\left(\frac{r}{M + 1} |y \right) = \frac{r + 1}{M + 1} - \frac{r}{M + 1} = \frac{1}{M + 1}
$$

Probably useless:

$$
P(N_\text{less} + K = r | \tilde\theta, y) = P\left(\frac{r}{M + 1} < \rho_{\phi,f}(\tilde\theta|y) \leq \frac{r + 1}{M + 1}\right) ?? \\

\int_\Theta \mathrm{d}\tilde\theta \int_Y \mathrm{d}y \: P\left(\frac{r}{M + 1} < \rho_{\phi,f}(\tilde\theta|y) \leq \frac{r + 1}{M + 1}\right) = 
\int_\Theta \mathrm{d}\tilde\theta \int_Y \mathrm{d}y \left(q\left(\frac{r + 1}{M + 1} |y \right) - q\left(\frac{r}{M + 1} |y \right) \right) = \\
\int_\Theta \mathrm{d}\tilde\theta \int_Y \mathrm{d}y \: q\left(\frac{r + 1}{M + 1} |y \right) - \int_\Theta \mathrm{d}\tilde\theta \int_Y \mathrm{d}y \: q\left(\frac{r}{M + 1} |y \right) = \frac{r + 1}{M + 1} - \frac{r}{M + 1} = \frac{1}{M + 1}
$$

More generally:

$$
q(x|y) = \\
= \int \mathrm{d} \tilde\theta \: \pi_\text{post}(\tilde\theta | y) 
  \mathbb{I}\left[
    x > C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) +  
    D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)  
  \right] + \\ 
  + \int \mathrm{d} \tilde\theta \: \pi_\text{post}(\tilde\theta | y) 
  \mathbb{I}\left[
    C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) +  
     D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)   >
    x > C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)  \right] \frac{x - C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)}{D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)} = \\
= \int \mathrm{d} \tilde\theta \: \pi_\text{post}(\tilde\theta | y) 
  \mathbb{I}\left[
    x > C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)  
  \right] + \\ 
  + \int \mathrm{d} \tilde\theta \: \pi_\text{post}(\tilde\theta | y) 
  \mathbb{I}\left[
    C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) +  
     D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)   >
    x > C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)  \right] 
    \left(\frac{x - C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)}{D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)} - 1\right) = \\
    C_f \left(C^{-1}_{\phi,f}(x | y) | y \right) + \\
    + \int \mathrm{d} \tilde\theta \: \pi_\text{post}(\tilde\theta | y) 
  \mathbb{I}\left[
    C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right) +  
     D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)   >
    x > C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)  \right] 
    \left(\frac{x - C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)}{D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)} - 1\right) = \\
$$

Assuming both $C_{\phi, f}$ and $C_{f}$ are discrete (over natural numbers)?

$$
q_{\phi,f}(x|y) =  C_f \left(C^{-1}_{\phi,f}(x | y) | y \right) + \\
    + \int \mathrm{d} \tilde\theta \: \pi_\text{post}(\tilde\theta | y) 
  \mathbb{I}\left[
    C_{\phi,f} \left(f \left(\tilde\theta, y \right)  + 1 | y \right) >
    x > C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)  \right] 
    \left(\frac{x - C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)}{D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)} - 1\right) = \\
    C_f \left(C^{-1}_{\phi,f}(x | y) | y \right) + \\
    + \int \mathrm{d} \tilde\theta \: \pi_\text{post}(\tilde\theta | y) 
  \mathbb{I}\left[
    f \left(\tilde\theta, y \right)  + 1 > C^{-1}_f(x | y) >
    f \left(\tilde\theta, y \right) \right] 
    \left(\frac{x - C_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)}{D_{\phi,f} \left(f \left(\tilde\theta, y \right) | y \right)} - 1\right) = \\
$$



## Correct posterior passes continuous SBC

outline: $C_f(C^{-1}_f(x|y)|y)$ is for a given $y$ either identity or "blocky". The tiebraking exactly removes the blocks, so we have the stronger condition

$$
\phi = \pi_{post}: 
\forall y \in Y, x \in [0,1] : q_{\phi,f}(x|y) = x
$$

Presumably, if the above condition holds, than it also holds for any transform of $f$, even if the posterior is incorrect (TODO check).

## Continuous SBC imples sample SBC

By contradiction: if sample ranks are non-uniform, there has to be some $x: \int_Y \mathrm{d} y: q_{\phi,f}(x|y) \neq x$.

## Sample SBC for all $M$ implies continuous SBC

Prove that $q$ is the limit of fractional rank distribution as $M \to \infty$

## Non-monotonous functions - simple case

Take $\Theta = \mathbb{R}$ and $\text{id} : Y \times \mathbb R \to \mathbb{R}, \text{id}(y, \theta) = \theta$   $C_\text{id}(s|y)$ is invertible over the domain of $f$ for all $y$. Take any continuous, almost everywhere smooth bijection $u: [0,1] \to [0,1], u(0) = 0, u(1) = 1$ and disjoint $Y_1, Y_2 \subset Y$ (with non-zero mass).  

$$
a,b \in [0,1]\\
\frac{a}{b} = \frac{\int_{y \in Y_2} \mathrm{d}y \: \pi_\text{marg}(y)}{\int_{y \in Y_1} \mathrm{d}y \: \pi_\text{marg}(y)}
$$

$$
v(x) = a u(x) + (1-a)x\\
w(x) = b (2x - u(x)) + (1-b)x\\

t_1(\theta, y) =  C^{-1}_\text{id}(v(C_\text{id}(\theta|y)) | y)\\
t_2(\theta, y) =  C^{-1}_\text{id}(w(C_\text{id}(\theta|y)) | y)
$$

$$
C^{-1}_\text{id}(v_i(C_\text{id}(\theta|y)) | y) = x \\
v_i(C_\text{id}(\theta|y)) = C_\text{id}(x |y) \\
C_\text{id}(\theta|y) = v^{-1}(C_\text{id}(x |y)) \\
\theta = C^{-1}_\text{id}(v^{-1}(C_\text{id}(x |y)) | y) \\
$$

$t_{1,2}$ are invertible and we can define:

$$
\omega(\theta|y) = \begin{cases}
  \pi_\text{post}(t^{-1}_1(\theta,y)|y) \frac{\partial t^{-1}_1}{\partial \theta}(\theta, y) | y) & \text{if } y \in Y_1 \\
  \pi_\text{post}(t^{-1}_2(\theta,y)|y) \frac{\partial t^{-1}_2}{\partial \theta}(\theta, y) | y) & \text{if } y \in Y_2 \\
  \pi_\text{post}(\theta | y) & \text{otherwise}
\end{cases}
$$

now $\omega$ satisfies SBC for $\text{id}$.


$$
y \in Y_1: C_{\omega,\text{id}}(s|y) = \int_{\Theta}\mathrm{d}\theta\mathbb{\,I}\left[\theta \leq s\right]\omega\left(\theta | y\right) = \\
\int_{\Theta}\mathrm{d}\theta\mathbb{\,I}\left[\theta \leq s\right]\pi_\text{post}(t^{-1}_1(\theta,y)|y) \frac{\partial t^{-1}_1}{\partial \theta}(\theta, y) | y) = \\
= \int_{\Theta}\mathrm{d}\bar\theta\mathbb{\,I}\left[t_1(\theta,y) \leq s\right]\pi_\text{post}(\bar\theta,y) = C_{t_1}(s|y) \\


$$

$$
y \in Y_1: q_{\omega,\text{id}}(x | y) = C_\text{id} \left(C^{-1}_{\omega,\text{id}}(x | y) | y \right) =
  C_\text{id} \left(C^{-1}_{t_1}(x | y) | y \right) = 
  \int \mathrm{d}\theta \: \mathbb{I}\left[\theta< C^{-1}_{t_1}(x | y)\right]\pi_\text{post}(\theta|y) = \\
  = \int \mathrm{d}\theta \: \mathbb{I}\left[C_{t_1}(\theta |y)< x\right]\pi_\text{post}(\theta|y) = \\
  = \int \mathrm{d}\theta \: \mathbb{I}\left[\int \mathrm{d}\tilde\theta\: \mathbb{I}\left[t_1(\tilde\theta, y) < \theta \right] \pi_\text{post}(\tilde\theta|y) < x\right]\pi_\text{post}(\theta|y) = \\
  = \int \mathrm{d}\theta \: \mathbb{I}\left[\int \mathrm{d}\tilde\theta\: \mathbb{I}\left[\tilde\theta < t_1^{-1}(\theta, y) \right] \pi_\text{post}(\tilde\theta|y) < x\right]\pi_\text{post}(\theta|y) = \\
  = \int \mathrm{d}\theta \: \mathbb{I}\left[C_\text{id}(t_1^{-1}(\theta, y)|y) < x\right]\pi_\text{post}(\theta|y) = \\
  = \int \mathrm{d}\theta \: \mathbb{I}\left[\theta < t_1(C^{-1}_\text{id}(x|y)\right]\pi_\text{post}(\theta|y) = \\
  = C_\text{id}\left(t_1(C_\text{id}^{-1}(x|y), y) \right) = v(x)
  
$$

Analogously

$$
y \in Y_2: q_{\omega,\text{id}}(x | y)  = w(x)
$$
So we have 

$$
\int_Y \mathrm{d}y \: q_{\omega,\text{id}}(x | y) \pi_\text{marg}(y) = \\
\int_{Y_1} \mathrm{d}y \: q_{\omega,\text{id}}(x | y) \pi_\text{marg}(y) + \int_{Y_2} \mathrm{d}y \: q_{\omega,\text{id}}(x | y) \pi_\text{marg}(y) + \int_{Y \setminus Y_1 \cup Y_2} \mathrm{d}y \: q_{\omega,\text{id}}(x | y) \pi_\text{marg}(y) = \\
v(x) \int_{Y_1} \mathrm{d}y \: \pi_\text{marg}(y) + w(x)\int_{Y_2} \mathrm{d}y\pi_\text{marg}(y) + x \int_{Y \setminus Y_1 \cup Y_2} \mathrm{d}y \: \pi_\text{marg}(y) = \\
(a u(x) + (1-a)x)\int_{Y_1} \mathrm{d}y \: \pi_\text{marg}(y) + (b (2x - u(x)) + (1-b)x)\int_{Y_2} \mathrm{d}y\pi_\text{marg}(y) + x \int_{Y \setminus Y_1 \cup Y_2} \mathrm{d}y \: \pi_\text{marg}(y) =\\
u(x) a \int_{Y_1} \mathrm{d}y \: \pi_\text{marg}(y) + b(2x - u(x)) \int_{Y_2} \mathrm{d}y\pi_\text{marg}(y) + x \left( \int_{Y \setminus Y_1 \cup Y_2} \mathrm{d}y \: \pi_\text{marg}(y) + (1-a)\int_{Y_1} \mathrm{d}y \: \pi_\text{marg}(y) + (1-b)\int_{Y_2} \mathrm{d}y\pi_\text{marg}(y)\right) \\
$$

We note that from requirements on $a,b$ we have $a \int_{Y_1} \mathrm{d}y \: \pi_\text{marg}(y) = b \int_{Y_2} \mathrm{d}y \: \pi_\text{marg}(y)$ and $u(x) + u^{-1}(x)$ se we have

$$
\int_Y \mathrm{d}y \: q_{\omega,\text{id}}(x | y) \pi_\text{marg}(y) = \\
\left(u(x) + 2x - u(x)\right)\left( a \int_{Y_1} \mathrm{d}y \: \pi_\text{marg}(y)\right) + x \left(1 - 2a \int_{Y_1} \mathrm{d}y \: \pi_\text{marg}(y) \right) = x \\

$$

$$
u(x) + v(x) = 2x \\
v(x) = 2x - u(x)
$$

Take $\Theta = \mathbb{R}$ and $f,g: Y \times \mathbb R \to \mathbb{R}, h: \mathbb{R} \to \mathbb{R}$ such that $\forall y \in Y, \theta \in \mathbb{R}: g(\theta, y) = h(f(\theta, y)$.  $C_f(\theta|y), C_g(\theta|y)$ are invertible for all $y$. There exist bounds $l_1 \leq u_1 \leq $\Theta


Take $\Theta = \mathbb{R}$ and $f,g: Y \times \mathbb R \to \mathbb{R}, h: \mathbb{R} \to \mathbb{R}$ such that $\forall y \in Y, \theta \in \mathbb{R}: g(\theta, y) = h(f(\theta, y)$.  $C_f(\theta|y), C_g(\theta|y)$ are invertible for all $y$. There exist bounds $l_1 \leq u_1 \leq $\Theta

$disjoint $Y_1, Y_2 \subset Y$ and disjoint $\Theta_1, \Theta_2 \subset \mathbb{R}$


