---
title: "Multivariate normal example"
output: html_notebook
---


```{r setup, message=FALSE,warning=FALSE, results="hide"}
library(SBC)
library(ggplot2)
library(mvtnorm)

options(mc.cores = parallel::detectCores())

library(future)
plan(multisession)

cache_dir <- "./_SBC_cache"

if(!dir.exists(cache_dir)) {
  dir.create(cache_dir)
}

```


We are using the model:

$$
\mathbf{\mu} \sim MVN(0, \mathbf{\Sigma})\\
\mathbf{y}_1, \ldots, \mathbf{y}_N \sim MVN(\mathbf{\mu}, \mathbf{\Sigma})\\
\mathbf{\Sigma} = \left(\begin{matrix}
 1 & 0.8 \\
 0.8 & 1 \\
\end{matrix}\right)
$$

where $MVN$ is the multivariate normal distribution. In this case the posterior has analytical solution and should also be multivariate normal.

```{r}
set.seed(546852)

mvn_sigma <- matrix(c(1, 0.8,0.8,1), nrow = 2)

generator_func_correlated <- function(N, sigma = mvn_sigma) {
  mu <- rmvnorm(1, sigma = sigma)
  y <- rmvnorm(N, mean = mu, sigma = sigma)
  list(variables = list(mu = mu[1,]),
       generated = list(y = y))
}

N_sims <- 1000

ds <- generate_datasets(SBC_generator_function(generator_func_correlated, N = 3, sigma = mvn_sigma), N_sims)

```


```{r}
my_backend_mvn <- function(sampling_func, N_samples = 100, func_extra_args = list()) {
  structure(list(sampling_func = sampling_func, N_samples = N_samples, 
                 func_extra_args = func_extra_args), class = "my_backend_mvn")
}

SBC_fit.my_backend_mvn <- function(backend, generated, cores) {
  all_args <- c(list(y = generated$y, N_samples = backend$N_samples), backend$func_extra_args)
  res_raw <- do.call(backend$sampling_func, all_args)
  
  K <- ncol(generated$y)
  colnames(res_raw) <- paste0("mu[", 1:K, "]")
  posterior::as_draws_matrix(res_raw)
}

SBC_backend_iid_draws.my_backend_mvn <- function(backend) {
  TRUE
}

my_backend_mvn_globals = c("SBC_fit.my_backend_mvn",
                                    "SBC_backend_iid_draws.my_backend_mvn",
                                    "mvn_sigma")
```


```{r}
quants <- generated_quantities(`mu[1] + mu[2]` = mu[1] + mu[2], 
                               `mu[1] - mu[2]` = mu[1] - mu[2],
                               `mu[1] * mu[2]` = mu[1] * mu[2],
                                mvn_log_lik = sum(mvtnorm::dmvnorm(y, mean = mu, sigma = mvn_sigma, log = TRUE)),
                               `mvn_log_lik[1]` = mvtnorm::dmvnorm(y[1,], mean = mu, sigma = mvn_sigma, log = TRUE),
                               `mvn_log_lik[3]` = mvtnorm::dmvnorm(y[3,], mean = mu, sigma = mvn_sigma, log = TRUE))
```


# Correct posterior

Introducing $\bar{\mathbf{y}} = \frac{1}{N}\sum_{i = 1}^{N} \mathbf{y}_i$, the posterior is $MVN\left(\frac{N\bar{\mathbf{y}}}{n + 1}, \frac{1}{N + 1}\mathbf{\Sigma}\right)  $

```{r}
sampling_func_correct <- function(y, N_samples, prior_sigma = mvn_sigma) {
  K <- ncol(y)
  N <- nrow(y)
  
  ybar = colMeans(y)

  post_mean <- N * ybar / (N  + 1)
  post_sigma <- prior_sigma / (N + 1) 
  res_raw <- mvtnorm::rmvnorm(N_samples, mean = post_mean, sigma = post_sigma)
}
backend_correct <- my_backend_mvn(sampling_func_correct)

res_correct <- compute_SBC(ds, backend_correct, gen_quants = quants, 
                              globals = my_backend_mvn_globals,
                              cache_mode = "results",
                              cache_location = file.path(cache_dir, "mvn_correct"))
```

```{r}
plot_rank_hist(res_correct)
plot_ecdf_diff(res_correct)
```

# Prior only

```{r}
sampling_func_prior_only <- function(y, N_samples) {
  mvtnorm::rmvnorm(n = N_samples, sigma = mvn_sigma)
}

backend_prior_only <- my_backend_mvn(sampling_func_prior_only)

res_prior_only <- compute_SBC(ds, backend_prior_only, gen_quants = quants, 
                              globals = my_backend_mvn_globals,
                              cache_mode = "results",
                              cache_location = file.path(cache_dir, "mvn_prior_only"))
```

```{r}
plot_rank_hist(res_prior_only)
plot_ecdf_diff(res_prior_only)
```


# One missing data point

```{r}
sampling_func_one_missing <- function(y, N_samples) {
  sampling_func_correct(y[1:(nrow(y) - 1),], N_samples)
}

backend_one_missing <- my_backend_mvn(sampling_func_one_missing)

res_one_missing <- compute_SBC(ds, backend_one_missing, gen_quants = quants, 
                              globals = c(my_backend_mvn_globals, "sampling_func_correct"),
                              cache_mode = "results",
                              cache_location = file.path(cache_dir, "mvn_one_missing"))
```

```{r}
plot_rank_hist(res_one_missing)
plot_ecdf_diff(res_one_missing)
```


# Incorrect Correlations



Especially when the number of data points is small, the correlations in the prior should persist in the posterior. 

We however generate posterior samples from a set of independent normal distributions that happen to have the correct mean and standard deviation, just the correlation is missing.

```{r}
sampling_func_uncorr <- function(y, N_samples, prior_sigma = 1) {
  K <- ncol(y)
  N <- nrow(y)
  
  ybar = colMeans(y)

  res_raw <- matrix(nrow = N_samples, ncol = K)
  for(k in 1:K) {
    post_mean <- N * ybar[k] / (N  + 1)
    post_sd <- sqrt(1 / (N + 1)) * prior_sigma 
    res_raw[,k] <- rnorm(N_samples, mean = post_mean, sd = post_sd)
  }
  res_raw
}

backend_uncorr <- my_backend_mvn(sampling_func_uncorr)


res_uncorr <- compute_SBC(ds, backend_uncorr, 
                        globals = my_backend_mvn_globals,
                        gen_quants = quants,
                        cache_mode = "results", 
                        cache_location = file.path(cache_dir, "mvn_uncorr"))
```


Although the posterior is incorrect, the default univariate checks don't show any problem even with `r N_sims` simulations.

```{r results_uncorr}
plot_rank_hist(res_uncorr)
plot_ecdf_diff(res_uncorr)
```

We can however add derived quantities that depend on both elements of mu. We'll
try their sum, difference, product and the multivarite normal log likelihood

```{r}
res_corr_gq  <- compute_SBC(ds, backend_uncorr, keep_fits = FALSE,
                        globals = my_backend_mvn_globals,
                        gen_quants = quants,
                        cache_mode = "results", 
                        cache_location = file.path(cache_dir, "corr_gq"))
```

We see that all of the derived quantities show problems, but with different strength
of signal. We'll especially note that the log likelihood is once again a very good
choice, while sum is probably the worst of those tested.

```{r results_corr_gq}
plot_rank_hist(res_corr_gq)
plot_ecdf_diff(res_corr_gq)
```

# Non-monotonous transform

```{r}
sampling_func_non_mon <- function(y, N_samples, prior_sigma = mvn_sigma) {
  # Sample as if correct
  K <- ncol(y)
  N <- nrow(y)
  
  ybar = colMeans(y)

  post_mean <- N * ybar / (N  + 1)
  post_sigma <- prior_sigma / (N + 1) 
  res <- mvtnorm::rmvnorm(N_samples, mean = post_mean, sigma = post_sigma)

  # Modify
  for(k in 1:K) {
    res_k <- res[,k]
    
    uniform_q <- pnorm(res_k, post_mean[k], sqrt(post_sigma[k,k]))
    if(mean(y[,k]) > 0) {
      transformed_q <- dplyr::if_else(uniform_q < 0.5, 1.5 * uniform_q, 0.75 + (uniform_q - 0.5)*0.5)
    } else {
      transformed_q <- dplyr::if_else(uniform_q < 0.5, 0.5 * uniform_q, 0.25 + (uniform_q - 0.5)*1.5)
    }
    res_k <- qnorm(transformed_q, post_mean[k], sqrt(post_sigma[k,k]))
    
    res[,k] <- res_k
  }
  res
}

backend_non_mon <- my_backend_mvn(sampling_func_non_mon)

quants_non_mon <- generated_quantities(`mu[1] * mu[2]` = mu[1] * mu[2],
                                       `abs(mu[1])` = abs(mu[1]),
                                       `abs(mu[2])` = abs(mu[2]),
                                mvn_log_lik = sum(mvtnorm::dmvnorm(y, mean = mu, sigma = mvn_sigma, log = TRUE)))


res_non_mon <- compute_SBC(ds, backend_non_mon, gen_quants = quants_non_mon, 
                              globals = c(my_backend_mvn_globals, "sampling_func_correct"),
                              cache_mode = "results",
                              cache_location = file.path(cache_dir, "mvn_non_mon"))
```

```{r}
plot_rank_hist(res_non_mon)
plot_ecdf_diff(res_non_mon)
```

# Small changes compound

```{r}
K_changes <- 2
set.seed(5665525)
mvn_sigma_changes <- matrix(0.8, nrow = K_changes, ncol = K_changes)
diag(mvn_sigma_changes) <- 1

ds_changes <- generate_datasets(SBC_generator_function(generator_func_correlated, N = 3, sigma = mvn_sigma_changes), n_sims = 1000)

sampling_func_small_change <- function(y, N_samples, prior_sigma) {
  res_correct <- sampling_func_correct(y, N_samples, prior_sigma)
  K = nrow(prior_sigma)
  bias <- rnorm(K, mean = 0, sd = 0.3)
  res <- res_correct
  for(k in 1:K) {
    res[,k] <- res[,k] + bias[k]  
  }
  #res_correct * matrix(ifelse(runif(K * N_samples) < 0.1, 0, 1), nrow = N_samples, ncol = K)
  res
}

backend_small_change <- my_backend_mvn(sampling_func_small_change, func_extra_args = list(prior_sigma = mvn_sigma_changes))

quants_change <- generated_quantities(sum = sum(mu),
                                      sum_abs = sum(abs(mu)),
                                mvn_log_lik = sum(mvtnorm::dmvnorm(y, mean = mu, sigma = mvn_sigma_changes, log = TRUE)))

res_small_change <- compute_SBC(ds_changes, backend_small_change, gen_quants = quants_change, 
                              globals = c(my_backend_mvn_globals, "sampling_func_correct", "mvn_sigma_changes"),
                              cache_mode = "results",
                              cache_location = file.path(cache_dir, paste0("mvn_small_change_", K_changes)))
```

```{r}
plot_rank_hist(res_small_change)
plot_ecdf_diff(res_small_change)
```


```{r}
library(tidyverse)
devtools::load_all()
theme_set(cowplot::theme_cowplot())
res_small_change$stats %>% 
  group_by(variable) %>%
  mutate(history = compute_diff_history_norm(rank, unique(max_rank), p = 1)) %>%
  filter(sim_id >= 10) %>%
  ggplot(aes(x = sim_id, y = history, color = variable)) + geom_line() +
  scale_x_log10() +
  scale_y_continuous("Mean absolute ECDF difference")
```

```{r}
gamma_thresholds_df <- get_precomputed_gamma_thresholds(K = 100)
```


```{r}
res_small_change$stats %>% 
  group_by(variable) %>%
  mutate(log_gamma = compute_log_gamma_history(rank, unique(max_rank))) %>%
  filter(sim_id > 10, sim_id < 500) %>%
  inner_join(gamma_thresholds_df, by = c("sim_id" = "N_sims")) %>%
  ggplot(aes(x = sim_id, y = log_gamma - log_gamma_threshold, color = variable)) + 
  geom_hline(yintercept = 0, color = "black") +
  geom_line() +
  scale_y_continuous("Log Gamma - Threshold")
```

```{r}
res_prior_only$stats %>% 
  group_by(variable) %>%
  mutate(log_gamma = compute_log_gamma_history(rank, unique(max_rank))) %>%
  filter(sim_id > 5, sim_id < 50) %>%
  inner_join(gamma_thresholds_df, by = c("sim_id" = "N_sims")) %>%
  ggplot(aes(x = sim_id, y = log_gamma - log_gamma_threshold, color = variable)) + 
  geom_hline(yintercept = 0, color = "black") +
  geom_line() +
  scale_y_continuous("Log Gamma - Threshold")
```

