---
title: "Simple Bernoulli"
output: html_notebook
---


# Simulations for the simple Bernoulli experiment


```{r setup}
library(SBC)
library(tidyverse)
library(patchwork)

library(future)
plan(multisession)
```


Generate datasets + setup a set of different generated quantities.

I create a new simple class of SBC backends that just use one function to generate samples when $y = 0$ and another when $y = 1$.

```{r}
set.seed(1558655)

N_sims_simple_gq <- 1000
N_sims_simple_gq_large <- 10000
N_samples_simple_gq <- 100

variables_simple_gq <- runif(N_sims_simple_gq_large)
generated_simple_gq <- 
  purrr::map(variables_simple_gq, ~ list(y = rbinom(1, size = 1, .x)))

datasets_simple_gq_large <- SBC_datasets(variables = posterior::draws_matrix(theta = variables_simple_gq),
                               generated = generated_simple_gq)

datasets_simple_gq <- datasets_simple_gq_large[1 : N_sims_simple_gq]


my_backend_func <- function(func0, func1) {
  structure(list(func0 = func0, func1 = func1), class = "my_backend_func")
}

SBC_fit.my_backend_func <- function(backend, generated, cores) {
  if(generated$y == 0) {
    posterior::draws_matrix(theta = backend$func0())
  } else if (generated$y == 1) {
    posterior::draws_matrix(theta = backend$func1())
  } else {
    stop("Invalid")
  }
}

SBC_backend_iid_draws.my_backend_func <- function(backend) {
  TRUE
}

my_backend_func_invcdf <- function(tfunc0, tfunc1) {
  structure(list(tfunc0 = tfunc0, tfunc1 = tfunc1), class = "my_backend_func_invcdf")
}

SBC_fit.my_backend_func_invcdf <- function(backend, generated, cores) {
  if(generated$y == 0) {
    posterior::draws_matrix(theta = backend$tfunc0(runif(N_samples_simple_gq)))
  } else if (generated$y == 1) {
    posterior::draws_matrix(theta = backend$tfunc1(runif(N_samples_simple_gq)))
  } else {
    stop("Invalid")
  }
}

SBC_backend_iid_draws.my_backend_func_invcdf <- function(backend) {
  TRUE
}


my_globals <- c("SBC_fit.my_backend_func", "SBC_backend_iid_draws.my_backend_func", "SBC_fit.my_backend_func_invcdf",
                "SBC_backend_iid_draws.my_backend_func_invcdf", "N_samples_simple_gq")

gq_simple_gq <- generated_quantities(
   log_lik = dbinom(y, size = 1, prob = theta, log = TRUE),
   sq = (theta - 0.5) ^ 2,
   sin3_2 = sin(3/2 * pi * theta),
   abs = abs(theta - 1/2),
   saw = ifelse(theta < 1/2, theta, -1/2 + theta),
   #saw_irreg = ifelse(theta < 0.375, theta, -1/2 + theta),
   swap = ifelse(theta < 1/2, theta, theta - 1),
   #sin_inv = sin(1 / theta),
   #y_plus_theta = y + theta,
   saw_quad = ifelse(theta < 1/2, theta^2, -1/2 + theta^3),
   
   # CRPS was suggested, but seems not really useful
   # Following equation (8) at https://arxiv.org/pdf/2002.09578v1.pdf for CRPS
   # CRPS = (1 - dbinom(y, size = 1, prob = theta)) - 0.5 * dbinom(1, size = 2, prob = theta)
   )

```

## Correct posterior

Setup a backend using the correct analytic posterior - it passes SBC including all GQs

```{r}
backend_simple_gq_ok <- my_backend_func(
  func0 = rlang::as_function(~ rbeta(N_samples_simple_gq, 1, 2)),
  func1 = rlang::as_function(~ rbeta(N_samples_simple_gq, 2, 1)))


res_simple_gq_ok <- compute_SBC(datasets_simple_gq, backend_simple_gq_ok, keep_fits = FALSE,
                        gen_quants = gq_simple_gq, globals = my_globals)

plot_rank_hist(res_simple_gq_ok)
plot_ecdf_diff(res_simple_gq_ok)
```

## Flipped 0 and 1 outcomes

Now we flip the functions - this still satisfies the "data-averaged posterior = prior" condition but actually fails even the "vanilla" sbc (i.e. the `theta` subplot). Interestingly, the `sq` quantity is completely insensitive to this flipping, because it is symmetric to flips in `theta` around $\frac{1}{2}$.

```{r}
backend_simple_gq_flip <- my_backend_func(
  func0 = rlang::as_function(~ rbeta(N_samples_simple_gq, 2, 1)),
  func1 = rlang::as_function(~ rbeta(N_samples_simple_gq, 1, 2)))


res_simple_gq_flip <- compute_SBC(datasets_simple_gq, backend_simple_gq_flip, keep_fits = FALSE,
                        gen_quants = gq_simple_gq, globals = my_globals)

plot_rank_hist(res_simple_gq_flip)
plot_ecdf_diff(res_simple_gq_flip)


```

## Counterexamples for vanilla SBC

We now build a counterexample, that is slightly similar to the true posterior and is designed to pass SBC for the `theta` parameter. The counterexamples are most naturally expressed vie inverse CDFs, so we'll show those. This is how the inverse CDFs of the counterexample compare to the correct ones:

```{r}
transform_fun_test1_0 <- function(u) {
  ifelse(u < 3/4, (2/3) * u, 0.5 + (u - 0.75)*2)
}

transform_fun_test1_1 <- function(u) {
  ifelse(u < 3/4, (1/3) * sqrt(2) * sqrt(u * (3 + 2 * u)), sqrt(3 - 6*u + 4*u^2))
}

plot_transform_funs <- function(func0, func1, name) {
  u <- seq(from = 0, to = 1,length.out = 100)
  plot1 <- rbind(data.frame(y = 0, u = u, invphi = func0(u), type = name),
        data.frame(y = 1, u = u, invphi = func1(u), type = name),
        data.frame(y = 0, u = u, invphi = 1 - sqrt(1 - u), type = "Correct"),
        data.frame(y = 1, u = u, invphi = sqrt(u), type = "Correct")
        ) %>%
    ggplot(aes(x = u, y = invphi, color = type)) + geom_line(size = 2) + facet_wrap(~y, labeller = label_both) +
    scale_y_continuous("Inverse CDF of theta") +
    scale_x_continuous("Quantile")
  
  plot2 <- 
    rbind(data.frame(y = 0, u = u, invphi_diff = func0(u) - ( 1 - sqrt(1 - u)), type = name),
        data.frame(y = 1, u = u, invphi_diff = func1(u) - sqrt(u), type = name),
        crossing(y = c(0,1), u = u, invphi_diff = 0, type = "Correct")) %>%
    ggplot(aes(x = u, y = invphi_diff, color = type)) + geom_line(size = 2) + facet_wrap(~y, labeller = label_both) +
    scale_y_continuous("Diff. inverse CDF of theta") +
    scale_x_continuous("Quantile")
  
  plot1 / plot2
}

plot_transform_funs(transform_fun_test1_0, transform_fun_test1_1, "Counterexample theta 1")
```

Let's run SBC. As designed, no problem with `theta`, but many GQs signal problems.

```{r}
backend_simple_gq_test1 <- my_backend_func_invcdf(transform_fun_test1_0, transform_fun_test1_1)


res_simple_gq_test1 <- compute_SBC(datasets_simple_gq, backend_simple_gq_test1, keep_fits = FALSE,
                        gen_quants = gq_simple_gq, globals = my_globals)

plot_rank_hist(res_simple_gq_test1)
plot_ecdf_diff(res_simple_gq_test1)


```

Just to test that our counterexample-generating method works, let's try another one:

```{r}
tfunc_test2_0 <- function(u) {
  u^2
}

tfunc_test2_1 <- function(u) {
  sqrt(u * (2-2*u+u^3))
}

plot_transform_funs(tfunc_test2_0, tfunc_test2_1, "Counterexample theta 2")
```


```{r}
backend_simple_gq_test2 <- my_backend_func_invcdf(tfunc_test2_0, tfunc_test2_1)


res_simple_gq_test2 <- compute_SBC(datasets_simple_gq, backend_simple_gq_test2, keep_fits = FALSE,
                        gen_quants = gq_simple_gq, globals = my_globals)

plot_rank_hist(res_simple_gq_test2)
plot_ecdf_diff(res_simple_gq_test2)


```

## Linear combinations

```{r}
tfunc_test_lincomb_0 <- function(u) {
  dplyr::if_else(runif(length(u)) < 0.5,
    tfunc_test2_0(u),
    transform_fun_test1_0(u)
  )
}

tfunc_test_lincomb_1 <- function(u) {
  dplyr::if_else(runif(length(u)) < 0.5,
    tfunc_test2_1(u),
    transform_fun_test1_1(u)
  )
}

backend_simple_gq_test_lincomb <- my_backend_func_invcdf(tfunc_test_lincomb_0, tfunc_test_lincomb_1)


res_simple_gq_test_lincomb <- compute_SBC(datasets_simple_gq_large, backend_simple_gq_test_lincomb, keep_fits = FALSE,
                        gen_quants = gq_simple_gq, globals = c(my_globals,
                                                               "tfunc_test2_0","tfunc_test2_1", "transform_fun_test1_0","transform_fun_test1_1"))

plot_rank_hist(res_simple_gq_test_lincomb)
plot_ecdf_diff(res_simple_gq_test_lincomb)


```



## Counterexamples for theta and log_lik

In the previous two cases, the `log_lik` variable always showed discrepancies. 
We can however construct counterexamples that satisfy both `theta` and `log_lik`,
here is one:

```{r}
tfunc_test3_0 <- function(u) {
  ifelse(u < 0.5, u *(2 - sqrt(2)),
           1 - sqrt(2 * (u - 1) *(1 - sqrt(2) + (-3 + 2 * sqrt(2))*u))
           )
}

tfunc_test3_1 <- function(u) {
  ifelse(u < 0.5, sqrt(2 * u *(-1 + sqrt(2) + (3 - 2 * sqrt(2))*u)), 
           sqrt(-3 + 2 * sqrt(2) + 2 * u *(5 - 3 * sqrt(2) +
           (-3 + 2 * sqrt(2))*u))
           )
}

plot_transform_funs(tfunc_test3_0, tfunc_test3_1, "Counterexample theta/log_lik 1")
```

The SBC for both `theta` and `log_lik` passes. Some other quantities however do show the failure! Note that the inverse CDFs follow the correct value quite closely and we need to run more than 1000 simulations (2500) to see violations, but we see them. This shows the space of useful quantities is not exhausted by a univariate marginal distribution and the (log)likelihood and that non-monotonous transformation of the univariate marginal can provide additional power to SBC.

```{r}
backend_simple_gq_test3 <- my_backend_func_invcdf(tfunc_test3_0, tfunc_test3_1)

res_simple_gq_test3 <- compute_SBC(datasets_simple_gq_large[1:2500], backend_simple_gq_test3, keep_fits = FALSE,
                        gen_quants = gq_simple_gq, globals = my_globals)

plot_rank_hist(res_simple_gq_test3)
plot_ecdf_diff(res_simple_gq_test3)
```

Let's construct another counter example - this one with larger deviations from
the correct posterior.

```{r}
tfunc_test4_0 <- function(u) {
  ifelse(u < 0.5, u^2 * 2 *(2 - sqrt(2)),
           1 - 2 * abs(u - 1) * sqrt(-4 + 3 * sqrt(2) + (-6 + 4*sqrt(2)) * (u -2) * u) 
           )  
} 

tfunc_test4_1 <- function(u) {
  ifelse(u < 0.5, sqrt(2 * u *(1 + 2*u *(-2 + sqrt(2) + (6 - 4*sqrt(2)) * u^2))),
           sqrt(-17 + 12 * sqrt(2) + 2 * u * (41 - 28 * sqrt(2) + 2 * u * (-34 + 23 * sqrt(2) + (-6+4*sqrt(2)) * (u - 4) * u))))
}

plot_transform_funs(tfunc_test4_0, tfunc_test4_1, "Counterexample theta/log_lik 2")
```

Once again, `log_lik` and `theta` show no problems (since we constructed the counterexample to match them), but _all the other_ variables
show the discrepancy!

```{r}
backend_simple_gq_test4 <- my_backend_func_invcdf(tfunc_test4_0, tfunc_test4_1)

res_simple_gq_test4 <- compute_SBC(datasets_simple_gq, backend_simple_gq_test4, keep_fits = FALSE,
                        gen_quants = gq_simple_gq, globals = my_globals)

plot_rank_hist(res_simple_gq_test4)
plot_ecdf_diff(res_simple_gq_test4)
```


## Prior only

As noted previously, using just the prior as posterior passes SBC for `theta` and all GQs that depend only on `theta`, however the `log_lik` GQ comes to the rescue! 

```{r}
backend_simple_gq_prior<- my_backend_func(
  func0 = rlang::as_function(~ runif(N_samples_simple_gq)),
  func1 = rlang::as_function(~ runif(N_samples_simple_gq)))


res_simple_gq_prior <- compute_SBC(datasets_simple_gq, backend_simple_gq_prior, keep_fits = FALSE,
                        gen_quants = gq_simple_gq, globals = my_globals)

plot_rank_hist(res_simple_gq_prior)
plot_ecdf_diff(res_simple_gq_prior)
```

## Summary

Of the quantities we use, we can construct counterexamples that satisfy any two of them, but combining `log_lik` with `theta` and a non-monotonic transformation of `theta` such as `sin(3/2*pi*theta)` manages to uncover all problems. At this point, I am not sure if those three restrict the solutions to only contain the correct posterior, or if a range of possible behaviours is still possible.
